---
title: "Tables in our reports"
author: "MB"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, warning = F, echo = F, message = F}
library(vispear)
library(ggplot2)
library(scales)
library(dplyr)
```


## `datatable_enr` function

### Overview

```{r, eval = F}
datatable_enr(data,
              options = list(),
              sorting = FALSE,
              desc_column = NULL,
              tooltip = "",
              ...)
```

### Use cases

The new function `datatable_enr` is the extention of `DT` package. It provides fixed `vispear` layout, and also, it facilitates the process of creating good-looking tables. We present several examples.

Below you can see a simple case with only one parameter - `data`.

```{r}
datatable_enr(data = mtcars)
```
 
 </br>

You can add `sorting` option in columns, by default, this option is disabled.

```{r}
datatable_enr(mtcars, sorting = T)
```

</br>

You can also add description of the table by using `tooltip` argument.

```{r}
datatable_enr(mtcars, tooltip = "This is the description of table.")
```

Besides, by using `desc_column` argument, you can differentiate first columns from others. It is useful especially when your table consists of information about titles.

```{r}
datatable_enr(mtcars, desc_column = 1:3, tooltip = "This is the description of table.")
```

In original `datatable` function (`DT` package), you can add several arguments in `options` argument. In order to allow you to customize a table, you can add whatever arguments in `options` which you would include in the original `datatable` function.

```{r}
datatable_enr(mtcars,
              desc_column = 1:3, 
              extensions = "Scroller",
              tooltip = "This is the description of table.")
```

Additionally, `datatable_enr` function has `...` argument in which you can add whaever argument you would include in the original `datatable` function, for example, `extensions` argument.

```{r}
mtcars %>% 
  bind_cols(mtcars) %>% 
  head(5) %>% 
  datatable_enr(tooltip = "This is a description.",
              options = list(fixedColumns = list(leftColumns = 3)),
              desc_column = 1:3,
              extensions = c('Scroller','FixedColumns'))
```

## `get_html_data` + `datatable_enr` function

### Overview

```{r, eval = F}
get_html_data(data, 
              column = NULL, 
              direction = 'col', #possible options: 'col', 'row', and 'area'
              formatter = 'simple_formatter') # any formatter you build
```

<br>

### Use cases

The `get_html_data` function (based on [formattable](https://renkun.me/formattable/) package) produces dataset with html tags which then, allows you to color cells within tables. Using with `datatable_enr`, it generates table you can include in your report. Below, the simple example of combining two functions is created. `mtcars` dataset is an input and three columns `(4:6)` were modified to add color cells. `direction` argument allows you to decide in which way the cells will be coloured. `area` option indicates that all 
values from all chosen columns will be taken into account. `col` option colors cells in each column separately, likewise `row` in each row. By default, the function uses `simple_formatter` which colors numeric values, meaning, the bigger value, the darker cell. Of course, you case build your own `formatter` (more details in formattable's [documentation](https://cran.r-project.org/web/packages/formattable/formattable.pdf)).

```{r}
get_html_data(mtcars, 
              column = 4:6, 
              direction = 'area') %>% 
  datatable_enr(tooltip = "This is a description.")
```

<br>

Additionally, `formatter` argument can changed using built-in formatter or whatever formatter you build. It depends on your needs. You can also add `column = -1` in order to
include all columns exept the first one.

```{r}
df <- cbind(data.frame(id = 1:5),
            do.call(cbind, lapply(1:8, function(x) sample(-50:50, 5))))


get_html_data(df, 
              column = -1, 
              direction = 'area', 
              formatter = 'sign_formatter') %>% 
  datatable_enr(tooltip = "This is a description.")
```

<br>

It is another example of using `direction = 'row'`.

```{r}
df <- cbind(data.frame(id = 1:5),
            do.call(cbind, lapply(1:8, function(x) sample(1:100, 5))))


get_html_data(df, 
              column = -1, 
              direction = 'row', 
              formatter = 'simple_formatter') %>%
  datatable_enr()

```

<br>

By default, `get_html_data` function uses one `formatter`, for all chosen columns (`column`).
However, it is possible to use as many formatters as you want. All you have to do, is to add a vector or list of formatters. Be careful, the number of columns and number of formatters have to be same.

```{r}
df <- data.frame(id = 1:5,
            simple_format = sample(-50:50, 5),
            sign_format = sample(1:100, 5))


get_html_data(df, 
              column = 2:3, 
              direction = 'col', 
              formatter = c('sign_formatter', 'simple_formatter')) %>%
  datatable_enr(tooltip = "This is a description.")

```

<br>

If you want to include percent values, it is critical to use `formattable::percent` in original dataset first.

```{r}
df <- data.frame(id = 1:5,
                 x = runif(5) %>% formattable::percent(0),
                 y = runif(5) %>% formattable::percent(0))
          

df %>% 
  get_html_data(column = -1) %>% 
  datatable_enr()
```

<br>

You can also use more options from `datatable_enr` if necessary.

```{r}
mtcars[,1:8] %>% 
  tibble::rownames_to_column() %>% 
  get_html_data(column = -1, direction = 'col') %>% 
  datatable_enr(tooltip = "This is a description.", desc_column = 1, 
              extensions = "Scroller",
              options = list(scrollY = 300))
```


### Warnings & Errors

```{r}
get_html_data(mtcars) %>%
  datatable_enr()
```

```{r, error = TRUE}
get_html_data(diamonds %>% head(),
                   column = 1:2,
                   direction = 'area') %>%
  datatable_enr()
```


```{r, error = TRUE}
get_html_data(mtcars,
                   column = 1:2,
                   direction = 'xxx') %>%
  datatable_enr()
```

